Deployable Probabilistic Programming

### 2019-04-16 paper/paper.tex
in progress

	There are two polar views on the role of probabilistic
	programming.  One view is that probabilistic programming is
	a flexible framework for specification and analysis of
	statistical models.  Proponents of this view perceive
	probabilistic programming as a tool for data science
	practitioners. A typical work flow consists of data
	acquisition and pre-processing, followed by several
	iterations of exploratory model design and testing of
	inference algorithms. Once a sufficiently robust statistical
	model is obtained, analysis results are post-processed,
	visualized, and occasionally integrated into algorithms of a
	production software system.

	The other, emerging, view is that probabilistic programming
	is an extension to a regular programming toolbox allowing
	algorithms implemented in general programming languages to
	have learnable parameters. In this view, statistical models
	are integrated into the software system, and inference and
	optimization take place during data processing, prediction,
	and algorithmic decision making in production.
	Probabilistic programming code runs unattended, and
	inference results are an integral part of the algorithms.

@@

	\begin{itemize}
		\item Guidelines for design and implementation of a
			probabilistic programming facility deployable as a part 
			of production software systems.
		\item A reference implementation of deployable probabilistic
			programming facility for the Go programming language.
		\item Design and implementation highlights outlining
			important choices we made in implementation of the
			probabilistic programming facility, and providing a
			basis for implementation of a similar facility for
			other general programming languages.
	\end{itemize}

@@

\subsection{Probabilistic ``Hello, World!''}

@@

A model is defined in its own package. The model must
implement interface \lstinline{Model} containing a single method
\lstinline{Observe}. \lstinline{Observe} accepts receives a slice of
\lstinline{float64} --- the model parameters --- and returns a
\lstinline{float64} scalar. For probabilistic inference, the
returned value is interpreted as the log-likelihood of the model
parameters. Inference models rely on computing the gradient of
the returned value with respect to the model parameters through
automatic differentiation. In the model's source code:

>>>>

	There are two polar views on the role of probabilistic
	programming.  One view is that probabilistic programming is
	a flexible framework for specification and analysis of
	statistical models.  Proponents of this view perceive
	probabilistic programming as a tool for data science
	practitioners. A typical workflow consists of data
	acquisition and pre-processing, followed by several
	iterations of exploratory model design and testing of
	inference algorithms. Once a sufficiently robust statistical
	model is obtained, analysis results are post-processed,
	visualized, and occasionally integrated into algorithms of a
	production software system.

	The other, emerging, view is that probabilistic programming
	is an extension of a regular programming toolbox allowing
	algorithms implemented in general programming languages to
	have learnable parameters. In this view, statistical models
	are integrated into the software system, and inference and
	optimization take place during data processing, prediction,
	and algorithmic decision making in production.
	Probabilistic programming code runs unattended, and
	inference results are an integral part of the algorithms.

@@

	\begin{itemize}
		\item Guidelines for design and implementation of a
			probabilistic programming facility deployable as a part 
			of a production software system.
		\item A reference implementation of deployable probabilistic
			programming facility for the Go programming language.
		\item Design and implementation highlights outlining
			important choices we made in implementation of the
			probabilistic programming facility, and providing a
			basis for implementation of a similar facility for
			other general programming languages.
	\end{itemize}

@@

\subsection{A Basic Example}

@@

A model is defined in its own package. The model must
implement interface \lstinline{Model} containing a single method
\lstinline{Observe}. \lstinline{Observe} accepts a slice of
\lstinline{float64} --- the model parameters --- and returns a
\lstinline{float64} scalar. For probabilistic inference, the
returned value is interpreted as the log-likelihood of the model
parameters. Inference models rely on computing the gradient of
the returned value with respect to the model parameters through
automatic differentiation. In the model's source code:
